Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-02-14 alloscore_manuscript/alloscore-application-shorter.Rnw
various edits, rephrasing

Suppose that a decision maker is tasked with determining how to allocate $K$ available units of a resource across $N$
locations. If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution
$F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes
the expected total unmet need according to the forecast. We will give a more precise mathematical statement in section
\ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is
\begin{align}
\sum_{i=1}^N \mathbb{E}_{F_i}[\text{unmet need in location $i$}], \label{eqn:informal_objective}
\end{align}
where the unmet need in a particular location is the difference between resource need in that location and the number of
resources that were allocated there. This allocation problem has an intuitively appealing solution: allocate so that the
probabilities of need exceeding allocation in various locations are as close to each other as possible. This will lead
to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single}
probability level $\tau$ that is shared in common for all locations.

After time passes and the actual level of resource need has been observed, the quality of a selected allocation can be
measured by comparing the actual need in each location to the amount of resources that were sent there. Specifically, we
compute the total unmet need that resulted from the selected allocation:
\begin{align}
    \sum_{i=1}^N \text{unmet need in location $i$}. \label{eqn:informal_loss}
\end{align}
%We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
One allocation is better than another if it results in lower total unmet need, and one forecast is better than another
if the allocation derived from it results in lower total unmet need.

The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation
that minimizes the expected unmet need according to that forecast. By ``avoidable unmet need'', we mean that the
allocation score does not include the amount of unmet need that was inevitable simply because the amount of available
resources $K$ was less than the need for resources. Rather, the allocation score measures the unmet need that could have
been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so
that nothing is wasted in one location while it could be put to use in another. An allocation score of 0 is optimal, and
indicates that no other allocation of resources could have met need better than the allocation suggested by $F$. A
larger allocation score indicates that it would have been possible to improve upon the allocation suggested by $F$.

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see section \ref{sec:bayes-quantiles} of the supplemental materials). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 unit of resources in location 1 and 4 units of resources
in location 2. If, on the other hand, $K = 10$ units are available, we will allocate 2 units of resources to location 1
and 8 units to location 2. Figure~\ref{fig:exp_alloc_example} illustrates the situation.

@@

Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
needs, we can measure the quality of the allocation suggested by the forecast by calculating the amount of unmet need
that resulted from that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ units
of the resource, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6
units of need unmet in location 2. However, working within the resource constraint, no other allocation could have done
better: for example, allocating 0 units of resources to location 1 and 5 to location 2 still results in a total unmet
need of 6 across both locations. Therefore, the forecast's allocation score is 0 with $K = 5$. On the other hand, when
$K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need
unmet in location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from
going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to
location 2. The allocation score for the forecast when $K = 10$ would therefore be 1 (= 2 realized $-$ 1 unavoidable) in
units of avoidable unmet need.

These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource
constraint is close to the observed need, because this is when it matters most which locations are allocated more or
less resources. If the amount of available resources is small relative to the observed need, any allocation of those
limited resources will result in a large amount of unmet need. If the amount of available resources is comparatively
large, it becomes less important which locations receive relatively more or fewer resources because all locations will
receive enough resources to meet their need. In either of these extremes of resource availability, the avoidable unmet
need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be
small.

@@

%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure the value of a forecast as an input to decision making. We keep the discussion here at a somewhat informal level; we refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger and Machina 2006; Ehm et al. 2016] for more technically precise discussion.

In the framework of decision theory, a decision corresponds to the selection of an action $x$ from some set of possible actions $\mathcal{X}$. For example, $x$ may correspond to the level of investment in a measure designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of all possible levels of investment that we might select. The quality of a decision to take a particular action $x$ is measured in relation to an outcome $y$ that is unknown at the time the decision is made. For example, $y$ may correspond to the number of individuals who eventually become sick and would benefit from the mitigation measure, and informally, an action $x$ is successful to the extent that it meets the realized need. In the face of uncertainty, a decision maker may use a forecast $F$ of the random variable $Y$ to help inform the selection of the action to take. We measure the value of a forecast as an input to this decision making process by the quality of the decisions that it leads to.

We can formalize the preceding discussion with the following three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
\item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$ eventually occurs.
\item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under the distribution $F$.
\item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section \ref{sec:proper} of the
supplement we demonstrate that Bayes scoring rules are proper by construction (a result also shown by \cite{dawid2007geometry,gneiting2007strictly}).

@@

In the decision making setting that we consider, an action $x = (x_1, \ldots, x_N)$ is a vector specifying the amount that is allocated to each of $N$ locations. We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$. The set $\mathcal{X}$ consists of all possible allocations that satisfy these constraints. The eventually realized resource need in each location is denoted by $y = (y_1, \ldots, y_N)$. These levels of need are not known at the time of decision making, so we define the random vector $Y = (Y_1, \ldots, Y_N)$ where $Y_i$ represents the as-yet-unknown level of resource need in location $i$. Forecasts of need in each location are collected in $F = (F_1, \ldots, F_N)$. We assume that resource need is non-negative and the forecasts reflect that, i.e. the support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that each unit of unmet need incurs a loss denoted by $L$, so that if the selected resource level $x_i$ in location $i$ is less than the realized need $y_i$, a loss of $L \cdot (y_i - x_i)$ results. A variety of extensions to this setup are possible; for example, we might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each location, or account for resource transportation costs. In this work, we choose to keep the loss function relatively simple to focus on the core ideas.

It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our setup assumes that $t_d < t_r$. Additionally, the structure of our loss captures a setting where the resource in question does not impact the amount of demand that will materialize at time $t_r$, but rather it is a resource that satisfies that demand. In the context of infectious disease, this means that we do not consider resources that are intended to reduce the number of people who will become sick at some point in the future, such as a preventative influenza or COVID-19 vaccine. Instead, our setup addresses resources like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time $t_r$; we do not explicitly consider sequences of multiple decisions that are made over time or account for the impact of decisions on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex decision making settings in the discussion.

With this problem formulation in place, we can develop a proper scoring rule following the outline in section \ref{sec:methods.detailed.decisiontheory}.

\paragraph{Step 1: specify a loss function.} The loss associated with a particular allocation is calculated by summing contributions from unmet need in each location:
\begin{equation}
s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i). \label{eqn:loss_fn}
\end{equation}
Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$ in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to unit $i$ is greater than or equal to the realized need. Also, $L$ is a constant scalar value, the same across all locations, specifying the ``cost'' of one unit of unmet need.

@@

The allocation score $S_A$ that we developed in the previous section measures the skill of the forecast distributions $F$ based on a single probability level $\tau^{F,K}$. This is appropriate if the resource constraint $K$ is a known constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the value of forecasts across a range of decision making scenarios with different resource constraints, we can use an \emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a distribution $p$:
$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
We note that the device of considering a range of hypothetical decision makers or decision making problems with different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.

@@

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS requires that the full cumulative distribution functions of the forecast distributions for each location are available. For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see the supplement for further details). In the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper. We use the procedure here to illustrate the properties of the score, and note that a forecaster or collaborative forecasting exercise interested in using the allocation score for evaluation could circumvent issues with propriety through thoughtful elicitation of representations of forecast distributions or by collecting the allocations at specified resource levels as part of forecast submissions.

@@

<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The x-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the per-capital allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = c(.5,1.5)) +
  ylab("Allocation score") +
  geom_point() + geom_line(aes(linewidth = as.factor(model == "per-capita")))+
  theme(legend.pos="none") +
  xlab(NULL)

@@

<<metrics-correlation, fig.width=8, fig.height=9, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.">>=
all_scores |>
  filter(as.Date(reference_date.x) < as.Date("2022-02-01"),
         model != "per-capita") |>
  dplyr::mutate(target_end_date = as.Date(target_end_date)) |>
  ggplot(aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
    geom_point(alpha=.7) +
    scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
    scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
#    scale_color_gradient("Target date") +
    scale_color_date(
      name="Target date",
      breaks = seq.Date(from=as.Date("2021-12-13"), by = 14, length.out=5)) +
    scale_size_continuous("Observed\nhospitalizations") +
    facet_wrap(.~model, ncol=3) +
    geom_abline(slope=1, intercept=0, linetype=2, color="gray") #+
#    theme(legend.pos="bottom")

@@

<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of K between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS). The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (light orange triangles) uses a uniform weight for all possible values of K. Note that the AS used in earlier sections of the application uses the single fixed value of K=15,000.  Panel C shows how either method of computing the IAS (y-axis) is correlated with AS at K=15,000 (x-axis). Every point represents the AS and IAS for one model. The IAS centered at 15,000 are more closely correlated with the AS values.">>=

@@

p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, color=IAS_type, shape=IAS_type)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, linetype=2) +
  scale_color_manual(name = "Integrated allocation score weighting", values = c("#386cb0", "#fdc086")) +
  scale_shape_manual(name = "Integrated allocation score weighting", values = c(1, 2)) +
  geom_abline(slope=1, intercept=0, color="gray") +
  xlab("Allocation score") +
  ylab("Integrated allocation score") +
  theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
        legend.background=element_rect(fill = alpha("white", 0.2)))

p2 <- slim_alloscore_2021.12.20 |>
  group_by(model, K) |>
  summarize(`Allocation score` = sum(components)) |>
  left_join(kprobs) |>
  ggplot(aes(x=K)) +
  geom_point(aes(y=weight), color = "#386cb0") +
  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
  geom_vline(xintercept = 15000, color="gray") +
  ylab("Weight")

p3 <- slim_alloscore_2021.12.20 |>
  group_by(model, K) |>
  summarize(`Allocation score` = sum(components)) |>
  mutate(model = factor(model, levels = rev(ias$model))) |>
  ggplot(aes(x=K, y=`Allocation score`, color=model, linetype=model)) +
  geom_line() +
  theme(legend.position = c(0.01,.99), legend.justification = c(0,1),
        legend.background=element_rect(fill = alpha("white", 0.2)),
        legend.title = element_blank(), legend.text = element_text(size=7))

>>>>

Suppose that a decision maker is tasked with determining how to allocate $K$ available units of a resource across $N$
locations. If the decision maker is provided with a multivariate forecast $F$ where each marginal forecast distribution
$F_i$ predicts resource need in a particular location, one option is to choose the resource allocation that minimizes
the expected total unmet need according to the forecast. We will give a more precise mathematical statement in section
\ref{sec:methods.detailed}, but informally, the total expected unmet need according to the forecast is
\begin{align}
\sum_{i=1}^N \mathbb{E}_{F_i}[\text{unmet need in location $i$}], \label{eqn:informal_objective}
\end{align}
where the unmet need in a particular location is the difference between resource need in that location and the number of
resource units that were allocated there. This allocation problem has an intuitively appealing solution: allocate so that the
probabilities of need exceeding allocation in various locations are as close to each other as possible. This will lead
to the allocations provided by $F$ being quantiles of the marginal distributions $F_i$ for some \emph{single}
probability level $\tau$ that is shared in common for all locations.

After time passes and the actual level of resource need has been observed, the value of a selected allocation can be
measured by comparing the actual need in each location to the amount of resources that were sent there. Specifically, we
compute the total unmet need that resulted from the selected allocation:
\begin{align}
    \sum_{i=1}^N \text{unmet need in location $i$}. \label{eqn:informal_loss}
\end{align}
%We emphasize that in Equation \eqref{eqn:informal_loss} the calculation of unmet need is based on the actual resource need that was realized in each location, while in Equation \eqref{eqn:informal_objective} the calculation of unmet need was based on the forecast distribution of future levels of resource need.
We regard one allocation as better than another (with respect to the realized need) if it results in lower total 
unmet need, and accordingly regard one forecast as better than another (again with respect to the realized need)
if the allocation derived from the former results in lower total unmet need than does the allocation derived from the 
latter.

The \textbf{allocation score} of the forecast $F$ is the avoidable unmet need that results from using the allocation
that minimizes the expected unmet need according to that forecast. By ``avoidable unmet need'', we mean that the
allocation score does not include the amount of unmet need that was inevitable simply because the amount of available
resources $K$ was less than the need for resources. Rather, the allocation score measures the unmet need that could have
been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so
that nothing is wasted in one location while it could be put to use in another. The best and lowest possible allocation 
score is 0.  A positive allocation score indicates some way in which an oracle could have used perfect foreknowledge to
meet more of the need than did the allocation suggested by $F$.

\paragraph{Example 1} Suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Exp}(1 / \sigma_1)$
and $F_2 = \mathrm{Exp}(1 / \sigma_2)$, where $\sigma_1 = 1$ and $\sigma_2 = 4$. The means of these distributions are
given by the scale parameters $\sigma_i$. When the marginal forecasts are exponential distributions, it can be shown
that the optimal allocation divides the available resources among the locations proportionally to the scale parameters
$\sigma_i$ (see section \ref{sec:bayes-quantiles} of the supplemental materials). If $K = 5$ units of the resource are
available, the optimal allocation according to $F$ would be 1 and 4 resource units in locations 1 and 2, respectively. 
Increasing $K$ to $10$ changes this optimal allocation to 2 and 8. Figure~\ref{fig:exp_alloc_example} illustrates the 
situation.

@@

Next suppose that we observe resource needs of 1 and 10 in locations 1 and 2, respectively. Based on these observed
needs, we evaluate the allocation suggested by the forecast by calculating the amount of unmet need
that resulted from that allocation over and above what was unavoidable given the resource constraint. With $K = 5$ 
available resource units, the allocation based on the forecast exactly meets the observed need in location 1, but it leaves 6
units of need unmet in location 2. But these 6 units of unmet need are unaviodable under the constraint $K=5$: allocating 0 units of resources to location 1 and 5 to location 2, for example, still results in a total unmet
need of 6 across both locations. This gives an allocation score of 0 for the forecast when $K = 5$. On the other hand, when
$K = 10$, the forecast's allocation results in $10 - 8 = 2$ units of unmet need in location 2 despite leaving no need
unmet in location 1. In this case, the oracle would be able to prevent all but 1 of the total 11 units of need from
going unmet, for example by allocating 1 unit of resources to location 1 and the remaining 9 units of resources to
location 2. So for $K = 10$, the forecast gets an allocaton score of 1 (= 2 realized $-$ 1 unavoidable) in
units of avoidable unmet need.

These scores illustrate a general result: allocation scores for a forecast will tend to be larger when the resource
constraint is close to the observed need, because this is when it matters most which locations are allocated more or
less resources. If the amount of available resources is small relative to the observed need, any allocation of those
limited resources will result in a large amount of unmet need. If the amount of available resources is comparatively
large, it becomes less important which locations receive relatively more or fewer resources because all locations are
likely to
receive enough resources to meet their need. In either of these extremes of resource availability, the avoidable unmet
need that arises from the allocation suggested by a forecast (i.e., the forecast's allocation score) will tend to be
small.

@@

%In this section, we give an overview of the decision theoretic setup for developing proper scoring rules that measure
%the value of a forecast as an input to decision making. We keep the discussion here at a somewhat informal level; we
%refer the reader to [some subset of Brehmer and Gneiting; Grünwald and Dawid; Dawid; Granger and Pesaran 2000; Granger
%and Machina 2006; Ehm et al. 2016] for more technically precise discussion.

Decision theory is concerned with understanding how an \emph{action} $x$ both is and ought to be selected from a set of
possible actions $\mathcal{X}$ --- that is, how a decision is and ought to be made --- based on the anticipated
consequences of $x$ under future states. We represent such future states as a random variable $Y$ and refer to a
realization $y$ of $Y$ as an \emph{outcome}. In this paper $x$ will be related to the level of investment in a measure
designed to mitigate severe disease outcomes such as hospital beds, ventilators, medication, or medical staff, with
$\mathcal{X}$ being the set of all feasible levels of investment, while an outcome $y$ will correspond to the number of
individuals who eventually become sick and would benefit from the mitigation measure. A basic strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
\emph{loss}, measuring the success of $x$, which in our examples will be the extent to which $x$ meets the realized need
$y$.  One way that a decision maker can try to anticipate the consequence of an action $x$ is by using a forecast $F$ of
$Y$. The loss of the action $x$ chosen according to $F$ can, in turn, be intepreted as the value of $F$ as an input to
the decision making process under the realized outcome $y$ of $Y$.

We arrive through this discussion at a three-step procedure for developing scoring rules for probabilistic forecasts:
\begin{enumerate}
  \item Specify a \emph{loss function} $s(x, y)$ that measures the loss associated with taking action $x$ when outcome $y$
    eventually occurs.
  \item Given a probabilistic forecast $F$, determine the \emph{Bayes act} $x^F$ that minimizes the expected loss under
    the distribution $F$.
  \item The \emph{scoring rule} for $F$ calculates the score as the loss incurred when the Bayes act was used: 
    $S(F, y) = s(x^F, y)$.
\end{enumerate}
This is a general procedure that may be applied in settings where it is possible to specify a quantitative loss
function. We call a scoring rule obtained from this procedure a \emph{Bayes scoring rule} and in section
\ref{sec:proper} of the supplement we review the basic fact from the scoring rule literature (e.g.,
\cite{dawid2007geometry,gneiting2007strictly}) that Bayes scoring rules are proper by construction.

@@

In our setting an action is a vector $x = (x_1, \ldots, x_N)$ specifying the amount of resources allocated to each of
$N$ locations. We require that $0 \leq x$, i.e., that each $x_i$ is non-negative, and that the total allocation across
all locations equals the amount of available resources, $K$: $\sum_{i=1}^N x_i = K$. The set $\mathcal{X}$ consists of
all possible allocations that satisfy these constraints. The eventually realized resource need in each location is
denoted by $y = (y_1, \ldots, y_N)$. The future outcome that will determine the loss caused by choosing the allocation
$x$ is the vector of resource needs $y = (y_1, \ldots, y_N)$ for each location which is a realization (unknown at the
time of decision making) of the random vector $Y = (Y_1, \ldots, Y_N)$. A forecast of the random need $Y$ for each
location is written as $F = (F_1, \ldots, F_N)$. We assume need is non-negative and the forecasts reflect that, i.e. the
support of each $F_i$ is a subset of $\mathbb{R}^+$. Finally, we assume that any unmet unit of need in any location
incurs a fixed loss of $L>0$, so that if the selected resource level $x_i$ in location $i$ is less than the realized
need $y_i$, a loss of $L \cdot (y_i - x_i)$ results. A variety of extensions to this setup are possible; for example, we
might account for storage costs for resources that go unused, allow for a different loss per unit of unmet need in each
location, or account for resource transportation costs. In this work, we choose to keep the loss function relatively
simple to focus on the core ideas.

It is helpful to clearly distinguish between the time $t_d$ when a \emph{d}ecision is made about a public health
resource allocation and the time $t_r$ when \emph{r}esource needs that might be addressed by that allocation occur. Our
setup assumes that $t_d < t_r$ and that the resource in question can only meet the need realized at $t_r$, not prevent
it. That is, we stipulate that the random vector $Y$ of need is independent from the chosen allocation $x$. In the
context of infectious disease, this means that we do not consider resources, such as vaccines, that are intended to
reduce the number of people who will become sick at some point in the future. Instead, our setup addresses resources
like hospital beds, oxygen supply, or ventilators which are intended to meet the medical needs of patients who are
already sick. Additionally, our setup addresses decision-making that is related to resource needs only at the time
$t_r$; we do not consider sequences of multiple decisions that are made over time or account for the impact of decisions
on resource needs at any time other than $t_r$. We outline some opportunities to extend our work to more complex
decision making settings in the discussion.

With this problem formulation in place, we can develop a proper scoring rule following the procedure in section
\ref{sec:methods.detailed.decisiontheory}.

\paragraph{Step 1: specify a loss function.} The loss incurred by an allocation $x$ under an outcome $y$ is the sum of
contributions from unmet need in each location:
\begin{equation}
  s_A(x, y) = \sum_{i=1}^N L \cdot \max(0, y_i - x_i). \label{eqn:loss_fn}
\end{equation}
Here, $\max(0, y_i - x_i)$ is the unmet need in location $i$, which is given by $y_i - x_i$ if the realized need $y_i$
in location $i$ is greater than the amount $x_i$ allocated to that location, or $0$ if the amount $x_i$ allocated to
unit $i$ is greater than or equal to the realized need. Also, $L$ is a constant scalar value, the same across all
locations, specifying the ``cost'' of one unit of unmet need.

@@

The allocation score $S_A$ developed in the previous section evaluates the forecast distributions
$F$ based on a single probability level $\tau^{F,K}$. This is appropriate if the resource constraint $K$ is a known
constant. However, if $K$ is not precisely known at the time of decision making or there is interest in measuring the
value of forecasts across a range of decision making scenarios with different resource constraints, we can use an
\emph{integrated allocation score} (IAS) that integrates the allocation score across values of $K$, weighting by a
distribution $p$:
$$S_{IAS}(F, y) = \int S_A(F,y; K) p(K) \, dK$$
We note that the device of considering a range of hypothetical decision makers or decision making problems with
different problem parameters has been employed in the past \cite[e.g.,][]{murphy1993whatisagoodforecast}.

@@

As described above, predictions were submitted to the Forecast Hub in the form of a set of 23 quantiles of the predictive distribution. The WIS can be directly calculated from these quantiles. However, calculation of the AS requires that the full cumulative distribution functions of the forecast distributions for each location are available. For the purpose of this analysis, to calculate the allocation score we approximated the full cumulative distribution functions based on the provided quantiles. On the interior of the provided quantiles we used a monotonic cubic spline to interpolate the quantiles, and in the lower and upper tails we used normal distributions with parameters selected so as to match the two lowest and two highest quantiles (see section \ref{sec:distfromq} the supplement for further details). In 
section \ref{sec:propriety_of_parametric_approximation} of the supplement, we show that if this evaluation procedure had been specified prospectively, the resulting score would be proper \textemdash but a post hoc application of this procedure is improper. We use the procedure here to illustrate the properties of the AS, and note that a collaborative forecasting hub interested in using the AS for evaluation could circumvent issues with propriety by communicating the definition of the AS to forecasters and then collecting allocations at specified resource levels as part of forecast submissions.

@@

<<metrics-over-time, fig.height=8, fig.width=8, fig.cap="Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the per-capita allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.">>=
p2 <- ggplot(alloscore_data_summaries,
             aes(x=target_end_date, y=alloscore, color=model, shape=model)) +
  scale_x_date(limits=range(score_data_summaries$target_end_date)) +
  scale_shape_manual(values=c(15:18, 4, 6, 15:18, 4, 6, 15:18, 4, 6)) +
  scale_linewidth_manual(values = c(.5,1.5)) +
  ylab("Allocation score") +
  geom_point() + geom_line(aes(linewidth = as.factor(model == "per-capita")))+
  theme(legend.pos="none") +
  xlab(NULL)

@@

<<metrics-correlation, fig.width=8, fig.height=9, fig.cap="Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The $x$- and $y$-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.">>=
all_scores |>
  filter(as.Date(reference_date.x) < as.Date("2022-02-01"),
         model != "per-capita") |>
  dplyr::mutate(target_end_date = as.Date(target_end_date)) |>
  ggplot(aes(x=mwis_rank_std, y=allo_rank_std, color=target_end_date, size=yobs)) +
    geom_point(alpha=.7) +
    scale_x_continuous("MWIS standardized rank", limits=c(0,1)) +
    scale_y_continuous("Allocation score standardized rank", limits=c(0,1)) +
#    scale_color_gradient("Target date") +
    scale_color_date(
      name="Target date",
      breaks = seq.Date(from=as.Date("2021-12-13"), by = 14, length.out=5)) +
    scale_size_continuous("Observed\nhospitalizations") +
    facet_wrap(.~model, ncol=3) +
    geom_abline(slope=1, intercept=0, linetype=2, color="gray") #+
#    theme(legend.pos="bottom")

@@

<<multi-k, fig.height=10, fig.width = 8, fig.cap = "Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS). The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (light orange triangles) uses a uniform weight for all possible values of $K$. Note that the AS used in earlier sections of the application uses the single fixed value of $K=15,000$.  Panel C shows how either method of computing the IAS ($y$-axis) is correlated with AS at $K=15,000$ ($x$-axis). Every point represents the AS and IAS for one model. The IAS centered at 15,000 are more closely correlated with the AS values.">>=

@@

p1 <- ggplot(ias_to_plot, aes(x=AS, y=IAS, color=IAS_type, shape=IAS_type)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, linetype=2) +
  scale_color_manual(name = "Integrated allocation score weighting", values = c("#386cb0", "#fdc086")) +
  scale_shape_manual(name = "Integrated allocation score weighting", values = c(1, 2)) +
  geom_abline(slope=1, intercept=0, color="gray") +
  xlab(expression(paste('Allocation score at ', italic('K'), ' = 15,000'))) +
  ylab("Integrated allocation score") +
  theme(legend.position = c(0.05,.9), legend.justification = c(0,1),
        legend.background=element_rect(fill = alpha("white", 0.2)))

p2 <- slim_alloscore_2021.12.20 |>
  group_by(model, K) |>
  summarize(`Allocation score` = sum(components)) |>
  left_join(kprobs) |>
  ggplot(aes(x=K)) +
  geom_point(aes(y=weight), color = "#386cb0") +
  geom_point(y=1/length(unique(slim_alloscore_2021.12.20$K)), color="#fdc086", shape = 2) +
  geom_vline(xintercept = 15000, color="gray") +
  xlab(expression(paste(italic('K')))) +
  ylab("Weight")

p3 <- slim_alloscore_2021.12.20 |>
  group_by(model, K) |>
  summarize(`Allocation score` = sum(components)) |>
  mutate(model = factor(model, levels = rev(ias$model))) |>
  ggplot(aes(x=K, y=`Allocation score`, color=model, linetype=model)) +
  geom_line() +
  xlab(expression(paste(italic('K')))) +
  theme(legend.position = c(0.01,.99), legend.justification = c(0,1),
        legend.background=element_rect(fill = alpha("white", 0.2)),
        legend.title = element_blank(), legend.text = element_text(size=7))
